Presentation Title

### 2017-06-30 thesis/chapters/ch1-intro.tex
Better analysis

Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
Which, for the first time, gave us a simple banking system in the form of an distributed ledger.
It needs no central control but still incorruptable even if there are malicious parties that aim to undermine the system.
We call such a ledger a blockchain.

The primary innovations are (1) its consensus model, which prevents double spending,
and (2) its incentive mechanism that encourages anyone (with adaquate hardware) to participate in the network and keep it running.
The double spending problem can be seen as an inconsistency issue.
For example, $C$ has 5 units of currency in her account.
At some instant in time, $A$ sees that $C$ transferred 5 units to $A$ and $B$ sees that $C$ transferred 5 units to $B$.
Hence, the transaction is inconsistent with respect to the observer.
Bitcoin and many other blockchain systems solve the inconsistency problem with a consensus algorithm.
The goal of the algorithm is reach agreement on a set of values, e.g. transactions, which eliminates inconsistencies.
In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving (using brute force) a puzzle.
The first miner to solve it generates a block containing all the collected transactions.
The miner is also rewarded with new coins and transaction fees.
It is important to note that every block, that is the solution of the puzzle,
depends on the previous block.
Hence the name blockchain.
Further, honest miners always work on the largest chain.
Due to the difficulty of the puzzle,
it is unlikely for more than one blockchain to exist in the network for a long period of time.
Thus every party sees a consistent blockchain which solves the double spending problem.
Further, due to the reward mechanism,
people are incentivised to act as miners and keep the system running.

@@

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound.
This is due to the fact that blocks are fixed to 1 MB are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
The issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propogate through the network.
It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
A recent empirical study Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
But given the bandwidth and latency constraints,
it not possible to have more than 758 transactions per second.
Even worse, it may give some miner an unfair lead over others.
Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa~\cite{visa},
which is in the order of tens of thousands transactions per second.

@@

We describe and analyse the aforementioned idea in the remainder of this work.
We begin with the problem description in~\Cref{ch:problem}.
This is followed by a detailed formulation of the model in~\Cref{ch:model}. 
Next, we analyse the correctness and performance of our design in~\Cref{ch:analysis}, this is where we present our key theorems.
Implementation and experimental results are discussed in~\Cref{ch:implementation}.
Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Chapters \ref{ch:related} and \ref{ch:conclusion} respectively.

>>>>

Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
Which, for the first time, gave us a simple banking system in the form of a distributed ledger.
It needs no central control but still incorruptable even if there are malicious parties that aim to undermine the system.
We call such a ledger a blockchain.

The primary innovations are (1) its consensus model which prevents double spending,
and (2) its incentive mechanism that encourages anyone (with adaquate hardware) to participate in the network and keep it running.
The double spending problem can be seen as an inconsistency issue.
For example, $C$ has 5 units of currency in her account.
At some instant in time, $A$ sees that $C$ transferred 5 units to $A$ and $B$ sees that $C$ transferred 5 units to $B$.
Hence the transaction is inconsistent with respect to the observer.
Bitcoin and many other blockchain systems solve the inconsistency problem with a consensus algorithm.
The goal of the algorithm is reach agreement on a set of valid values, e.g. transactions. 
This effectively eliminates inconsistencies.
In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving (using brute force) a difficult puzzle.
The first miner to solve it generates a block containing all the collected transactions.
The miner is also rewarded with new coins and transaction fees.
It is important to note that every block, that is the solution of the puzzle,
depends on the previous block.
Hence the name blockchain.
Further, honest miners always work on the largest chain.
Due to the difficulty of the puzzle,
it is unlikely for more than one blockchain to exist in the network for a long period of time.
Thus every party sees a consistent blockchain which solves the double spending problem.
% Further, due to the reward mechanism,
% people are incentivised to act as miners and keep the system running.

@@

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound.
This is due to the fact that blocks are fixed to 1 MB are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propogate through the network.
It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
But given the bandwidth and latency constraints,
it not possible to have more than 758 transactions per second.
Even worse, it may give some miner an unfair lead over others.
Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa~\cite{visa},
which is in the order of tens of thousands transactions per second.

@@

We describe and analyse our idea in the remainder of this work.
We begin with the problem description in~\Cref{ch:problem}.
This is followed by a detailed formulation of the model in~\Cref{ch:model}. 
Next, we analyse the correctness and performance of our design in~\Cref{ch:analysis}, this is where we present our key theorems.
Implementation and experimental results are discussed in~\Cref{ch:implementation}.
Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Chapters \ref{ch:related} and \ref{ch:conclusion} respectively.

### 2017-06-30 thesis/chapters/ch4-analysis.tex
Better analysis

\subsection{Bandwidth Requirement for Transactions}
With our assumption on the time complexity of ACS,
we are ready to analyse the bandwidth.
We know that to create and then validate a transaction, the bandwidth required is of $O(l)$,
where $l$ is the length of the agreed fragment.
This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment,
the other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
Then
$$l = r_{\text{tx}} D_{\text{acs}},$$
where $D_{\text{acs}}$ is the duration an instance of ACS.
But from \Cref{sec:acs-time-complexity},
we know that $D_{\text{acs}}$ is of $O(N)$, thus the bandwidth per transaction is $O(N)$.
% TODO how does message complexity translate to time complexity?
% TODO computational task minimal compared to messages, thus the sending of message is expensive?
This is intuitive because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
This behaviour is also verified experimentally in \Cref{ch:implementation}.

\subsection{Global Throughput}
Using our results so far, we are able to analyse the global throughput.
Suppose every node has some fixed bandwidth capacity $C$ (unit of communication per second),
they make transactions at $r_{\text{tx}}$ per second.
Then we have the inequality $C \ge r_{\text{tx}} l$, where $l$ is the length of the the agree fragment as before.
Rearranging, we get
$$\frac{C}{l} \ge r_{\text{tx}}.$$
With this, we consider two cases, first is  when all the nodes are running at maximum capacity, i.e. $\frac{C}{l} = r_{\text{tx}}$.
Recall that $l$ is of $O(N)$, so as the the population increases, 
the transaction rate must decrease in order to maintain the inverse relationship,
thus $r_\text{tx}$ is of $O(N^{-1})$. % TODO is this formal? use omega?
Therefore, the global throughput would be $O(N^{-1})N = O(1)$.
In the second case, nodes are not running at maximum capacity and $r_{\text{tx}}$ is maintained.
Therefore, if every node runs at $r_{\text{tx}}$,
the global throughput becomes $O(N)$ until $N$ is too large and we go back to the first case.
A visualisation of the result is shown in \Cref{fig:throughput-visualisation}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{throughput-visualisation}
  \caption{Visualisation of the worst case global throughput.}
  \label{fig:throughput-visualisation}
\end{figure}

The upshot of this analysis is that our system scales nicely at a global throughput of $O(N)$ until the population gets too large.
Then we maintain a constant global throughput.
This result falls a bit short of what we envisioned in the introduction.
However, the population is not the number of nodes that use the system, but the number of nodes that are online during a single round.
Furthermore, this result is for the worst case where every transaction needs an agreed fragment to be transmitted.
In practice, nodes are able to cache agreed fragments.
For instance, if $u$ and $v$ make $x$ transactions in a single round,
then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.

@@

On the other hand, due to the fact that hypergeometric distributions have light tails,
that is ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when 
the expected value is  much smaller than $\lfloor \frac{n-1}{3} \rfloor$ is small.
For example, we can use tail inequality to bound the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when only 5\% are black.
The tail inequality is
$$
\Pr[X \ge E[X] + \tau n] \le e^{-2\tau^2n},
$$
where $E[X] = n/20$.
We are interested in $\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]$, so
\begin{align*}
\tau &= \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \frac{1}{20} \\
  &> \frac{\lfloor \frac{n-1}{3} \rfloor}{n} - \frac{1}{20} \\
  &\ge \frac{1}{4} - \frac{1}{20} = \frac{1}{5} \quad \text{for $n \ge 4$}.
\end{align*}
Putting $\tau$ back into the tail inequality we get
$$
\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-\frac{2}{25} n}.
$$
The bound is not tight, but it is clear that for sufficiently large $n$, the probability becomes vanishingly small.
Hence, it is possible to use our system in practice when 5\% of the population are adversarial but $n \ge 3t + 1$ does not hold,
as long as $n$ is large.

>>>>

\subsection{Data Volume Requirement for Transactions}
With our assumption on the time complexity of ACS,
we are ready to analyse the amount of data required to be transmitted over a link per transaction,
which we call the data volume per transaction.
% First, we define bandwidth as the ``amount of data a link or network path can deliver per unit of
% time''~\cite{prasad2003bandwidth}.
We know that to create and then validate a transaction,
the data volume per transaction is of $O(l)$,
where $l$ is the length of the agreed fragment.
This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment,
the other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
Then
$$l = r_{\text{tx}} D_{\text{acs}},$$
where $D_{\text{acs}}$ is the duration an instance of ACS.
But from \Cref{sec:acs-time-complexity},
we know that $D_{\text{acs}}$ is of $O(N)$, thus the data volume per transaction is $O(N)$.
This is intuitive because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
The behaviour is also verified experimentally in \Cref{ch:implementation}.

\subsection{Global Throughput}
Using our results so far, we are able to analyse the global throughput.
First we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
Now, suppose every node has some fixed bandwidth per link $C$ and $N$ links.
They also make transactions at $r_{\text{tx}}$ per second.
Then we have the inequality
$$NC \ge r_{\text{tx}} l,$$
where $l$ is the length of the the agree fragment as before.
The inequality suggests that the rate for which transactions and validations are made cannot exceed the total bandwidth of all the links.

We note that the inequality does not hold if the node is only transacting with a small sample of some subset of the population.
This is because it cannot use all the bandwidth available in all the links.
For instance, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages.
Thus multiple transactions can be validated with a single \texttt{vd\_resp} message.
For this work, we analyse the worst case where every node transacts with a random node from the population,
and a new \texttt{vd\_resp} must be sent for every \texttt{vd\_req} message.

Consider the case where the system is making use of all the bandwidth, i.e. $NC = r_{\text{tx}} l$.
Recall that $l$ is of $O(N)$, that means LHS and RHS both grow linearly with respect to $N$.
Hence, there exist some constant $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
Finally, if every node in the network is transacting at a constant rate, then the global throughput (in terms of transactions per second) is of $O(N)$.
If the system is not making use of all the bandwidth.
We also maintain a constant transaction rate by the same argument.
Thus a global throughput is also $O(N)$.
We verify both of these claims experimentally in~\Cref{sec:evaluation}.

% The upshot of this analysis is that our system scales nicely at a global throughput of $O(N)$ until the population gets too large.
% Then we maintain a constant global throughput.
% This result falls a bit short of what we envisioned in the introduction.
% However, the population is not the number of nodes that use the system, but the number of nodes that are online during a single round.
% Furthermore, this result is for the worst case where every transaction needs an agreed fragment to be transmitted.
% In practice, nodes are able to cache agreed fragments.
% For instance, if $u$ and $v$ make $x$ transactions in a single round,
% then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.

@@

On the other hand, due to the fact that hypergeometric distributions have light tails,
that is ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when 
the expected value is  much smaller than $\lfloor \frac{n-1}{3} \rfloor$ is small.
We can use tail inequality to bound the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when only $n\alpha$ are black where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor / n$.
The tail inequality is
$$
\Pr[X \ge E[X] + \tau n] \le e^{-2\tau^2n},
$$
where $E[X] = n\alpha$.
We are interested in $\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]$, so
\begin{align*}
\tau &= \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha
\end{align*}
Putting $\tau$ back into the tail inequality we get the following bound.
$$
\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-2 \big(\frac{\lfloor  \frac{n - 1}{3} \rfloor + 1}{n} - \alpha \big)^2 n}
$$

The bound is not tight, but it is useful for for picking parameters.
For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,
the probability is maximum when the squared term is minimum at $\alpha = \lfloor \frac{n-1}{3} \rfloor/n$.
The probability is minimum when the squared term is maximum at $\alpha = 0$.
Hence, if $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
To put this into perspective, suppose $n = 1000$ and $\alpha = 1/20$, then the probability to draw more black balls than the threshold is only $2 \times 10^{-48}$.
